\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage[margin=1.25in]{geometry}
%\setlength{\parindent}{0cm}

\title{Boundary Effects in Stochastic Cyclic Competition Models on a Two-Dimensional Lattice}
\author{M. Lazarus Arnau, Shannon R. Serrao, Uwe C. T{\"a}uber\\ Department of Physics (MC 0435) and Center for Soft Matter and Biological Physics\\ Virginia Tech, Blacksburg, Virginia 24061}
%\date{July 26, 2018}

\begin{document}
    
    \maketitle
    We study noise-induced and -stabilized spatial patterns in two distinct 
    stochastic population model variants for cyclic competition of three species, 
    namely the Rock-Paper-Scissors (RPS) and the May-Leonard (ML) models. In two
    dimensions, it is well established that the ML model can display 
    (quasi-)stable spiral structures, in contrast to simple species clustering 
    in the RPS system. Our ultimate goal is to develop local control schemes 
    which allow us to affect the formation of these spatio-temporal patterns. 
    To this end, we have employed MC simulations to investigate how changing 
    the microscopic rules in a subsection of a two-dimensional lattice 
    influences the macroscopic behavior in the rest of the lattice. 
    Specifically, we implement the ML reaction scheme on a torus, except on a 
    ring-shaped patch, which is set to follow the cyclic Lotka-Volterra 
    predation rules of the RPS model. At the RPS-ML interface we observe a 
    marked disruption of the usual spiral patterns in the form of plane waves 
    emanating from the RPS region. Also, we report a distinct decrease in 
    local population density near the interface in comparison to the bulk of the 
    ML region.

    %\hskip 

    %We will continue to characterize these effects before using our deepened understanding
    %of these systems to develop possible methods for local external control. We plan to
    %begin by exploring the possibility of using pulsed and periodic seeding to either
    %stabilize or destroy the spiral patterns displayed by our simulations.

    \hskip

    Research was sponsored by the U.S. Army Research Office and was accomplished 
    under Grant Number W911NF-17-1-0156. 
\end{document}
